---
title: "Lab 6"
author: "Bri Baker"
date: "November 9, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(ggpubr)
library(broom)
```

# Part 1: A rank-based test example (Mann Whitney U)

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)

```

```{r}
hist(gp_1)

hist(gp_2)
```

Not clearly normally distributed from exploratory histograms
Somewhat small sample size (n = 15 for each)
I’ve decided that ranks (or, medians) are a more valuable metric to compare for these data

```{r}
mymwu <- wilcox.test(gp_1, gp_2)

print(mymwu)
```



# Part 2: Simple linear regression

```{r}
#visualize
ggplot(data = penguins, aes(flipper_length_mm, body_mass_g)) +
  geom_point()



```



```{r}
#model
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

summary(penguin_lm)

```

### Access model outputs

We can access the coefficients for the model using:  

- The slope is `r round(penguin_lm$coefficient[2],2)` (g / mm)
- The y-intercept is `r round(penguin_lm$coefficient[1],2)` (g)
- The full equation is mass = `r round(penguin_lm$coefficient[2],2)`*(flipper length) + (`r round(penguin_lm$coefficient[1],2)`)

..this is tedious and there is a better way

```{r}
#model output information
penguin_lm_tidy <- broom::tidy(penguin_lm)


# Get the intercept: 
penguin_int <- penguin_lm_tidy$estimate[1]
penguin_int

# Then to get the flipper_length coefficient:
penguin_coef <- penguin_lm_tidy$estimate[2]
penguin_coef

#at a glance
penguin_lm_out <- broom::glance(penguin_lm)
penguin_lm_out
```

So now I can say:

"Simple linear regression was used to explore the relationship between penguin flipper length (mm) and body mass (g) across all three penguin species, and including both male and female penguins. A significant regression model was found ($\beta$ = `r round(penguin_coef,3)`, F(`r penguin_lm_out$df`,`r penguin_lm_out$df.residual`) = `r round(penguin_lm_out$statistic,1)`, p < 0.001) with an R^2^ of `r round(penguin_lm_out$r.squared,3)`."

### Explore model assumptions

The first one: fitted values vs. residuals
The second one: QQ-plot for residuals
The third one: another way of looking at fitted vs. residuals (these are just standardized residuals, but you can interpret it the same way)
The fourth one: Cook’s distance, a measure of “influence” or “leverage” that individual points have on the model - often considered a way to explore outliers.

```{r}
plot(penguin_lm)
```


### Visualize

Use `geom_smooth(method = "lm")` to add a linear model to an existing scatterplot

Use `stat_cor()` and/or `stat_regline_equation()` to add equation information directly to the plot panel, at an x- and y-position that you specify (and yes, you can mess with the digits & appearance here)

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 180, label.y = 5700)
```


### Find Pearson’s r for correlation

```{r}
penguins_cor <- cor.test(penguins$flipper_length_mm, penguins$body_mass_g)

print(penguins_cor)
```
Here, we see that there is a strong positive correlation between penguin flipper length and body mass (*r* = `r round(penguins_cor$estimate,2)`, t(`r penguins_cor$parameter`) = `r round(penguins_cor$statistic,2)`, p < 0.001). 
